Pool Model - A unified Mathematical\\ Modelling Framework for Food Science

### 2025-10-31 paper/pool_model.tex
tex: update abstract

\begin{abstract}
    Bacterial consortia are complex systems which are often comprised of multiple distinct species.
    Their dynamics on a population-based level often result from interactions that lead to
    interspecies cooperation or competition.
    Various empirical and inactivation models have been conceived and applied to problems related to
    food microbiology.
    However, the utilized models live on the population level, omitting any spatial effects such as
    heterogeneity or diffusion.
    In this work we present a novel generalized framework, called "pool model" which uses multiple
    distinct pools that can hold bacterial states of distinct species or extracellular nutrient
    abundance.
    It allows us to describe these systems mechanistically as variations of an underlying, more
    fundamental model.
    To illustrate the usefulness of the model, multiple different examples are reported.
    Finally, we construct an \acl{abm} for a given pool model, showcasing that the resulting model
    arises naturally from the individual behaviour of many cellular agents.
    We further apply this \ac{abm} to study the role of spatial effects on the dynamics of bacterial
    growth.
\end{abstract}

>>>>

\begin{abstract}
    Bacterial consortia are complex systems which are often comprised of multiple distinct species.
    Their dynamics on a population-based level often result from various intra- and inter-species interactions,
    including cooperation or competition.
    Various empirical and mechanistic models have been conceived and applied to problems related to food microbiology.
    In this work we present a generalized framework, which uses multiple distinct "pools" that can distinguish
    between microbial types in different metabolic states or describe the extracellular environment properties.
    It allows us to incorporate the underlying biological processes and construct the \ac{ode} models of
    different complexity levels.
    To illustrate the application scope of the model, multiple different examples of the microbial systems
    are reported.
    Finally, to explore the limitations of the \ac{ode} formulation, we analyze the influence of the spatial effects.
    For this, we construct an \acl{abm} for a two-species system, showcasing that the resulting pool model
    arises naturally from the individual behaviour of many cellular agents.
    We further apply this \ac{abm} to study the role of properties such as diffusion and spatial heterogeneity
    on the dynamics of bacterial growth.
\end{abstract}
